---
title: "8.1 Exercise"
author: "Joaquin Rodriguez"
date: "10/15/2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(ggfortify)
library(lmtest)
library(stargazer)
library(broom)
library(knitr)

abramowitz <- read.csv(file = "data/abramowitz.txt")
abramowitz <- 
  abramowitz %>%
  mutate(Term = as.factor(Term))
campbell <- read.csv(file = "data/campbell.txt")
holbrook <- read.csv(file = "data/holbrook.txt")
holbrook <- 
  holbrook %>% 
  mutate(Tenure = as.factor(Tenure))
lockerbie <- read.csv(file = "data/lockerbie.txt")
```

## Campbell Model
```{r}
campbell %>% 
  gather(-Year, key = "Var", value = "Value") %>% 
  ggplot(aes(Year, Value)) +
  geom_line() +
  facet_grid(Var ~ ., scales = "free")
```

```{r}
campbell %>% 
  gather(-Year, -IncVote, key = "Var", value = "Value") %>% 
  ggplot(aes(Value, IncVote)) +
  geom_point() +
  facet_grid(. ~ Var, scales = "free")
```


```{r}
lm.campbell <- lm(IncVote ~ SepTr + GDPGrR , data = campbell)
summary(lm.campbell)

autoplot(lm.campbell, which = 1:6, ncol = 2, label.size = 3)
```

```{r}
bgtest(lm.campbell, order = 5)
```


## Abramowitzs Model

```{r}
abramowitz %>%
  mutate(Term = as.integer(Term)) %>%
  gather(IncVote, Term, PresPop, GDPGrR, key = "Var", value = "Value") %>%
  ggplot(aes(Year, Value)) +
  geom_line() +
  facet_grid(Var ~ ., scales = "free")
```

```{r}
abramowitz %>%
  mutate(Term = as.integer(Term)) %>%
  gather(Term, PresPop, GDPGrR, key = "Var", value = "Value") %>% 
  ggplot(aes(Value, IncVote)) +
  geom_point() +
  facet_grid(. ~ Var, scales = "free")
```


```{r}
lm.abramowitz <- lm(IncVote ~ PresPop + GDPGrR, data = abramowitz)
summary(lm.abramowitz)

autoplot(lm.abramowitz, which = 1:6, ncol = 2, label.size = 3)
```

```{r}
# H0: there is no serial correlation up to p
bgtest(lm.campbell, order = 5)
```

## Holbrook Model
```{r}
lm.holbrook <- lm(IncVote ~ PresPop + PersFin + Tenure, data = holbrook)
lm.holbrook.summ <- summary(lm.holbrook)
```

The Holbrook model assumes that there the predictors of incumbent vote are:  
1. Presential popularity: average percentage of the public who said that they approved of the way the president is handling his job.  
2. Personal finance: measure of restrsopective personal finances based on responses to the Survey of Consumers.  
3. Tenure in office: binary variable coded 1 for presidential candidates whose party has held the White House for two or more terms and 0 for candidates whose party has held the White Whouse for one term.  

Following we observe the trend of the four indepedent variables over the analyzed period. 

```{r}
holbrook %>%
  mutate(Tenure = as.integer(Tenure)) %>% 
  gather(IncVote, PresPop, PersFin, Tenure, key = "Var", value = "Value") %>% 
  ggplot(aes(Year, Value)) +
  geom_line() +
  facet_grid(Var ~ ., scales = "free") +
  labs(title = "Variables of interest over the years")
```

Since we are dealing with we perform the Breusch-Godfrey Test for autocorrelation up to the firth order. 

```{r}
bgtest(lm.holbrook, order = 5)
```

The test indicates that there is no autocorrelation of the error terms, therefore the assumption of the linear model is satisfied.  

Following we observe the correlation between the incumbent vote with the presidential popularity and personal finances measures.

```{r results="asis"}
stargazer(cor(holbrook[, 2:4]), type = "html")
```

We observe how the incumbent vote is highly correlated to the presidential popularity and personal finances measures. Furthermore, we observe how the presidential popularity and personal finances measures have also a high correlation, which might produce multicollinearity problems in the model.  

Following we investigate the relationship between the analyzed measures.  

```{r}
holbrook %>%
  gather(PresPop, PersFin, key = "Var", value = "Value") %>% 
  ggplot(aes(Value, IncVote)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "lm", se = T) +
  facet_grid(. ~ Var, scales = "free") +
  labs(title= "Relationship between Incumbent Vote and Presidential Popularity / Personal Finances")

holbrook %>%
  ggplot(aes(Tenure, IncVote)) +
  geom_boxplot() +
  labs(title= "Relationship between Incumbent Vote and Tenure")
```

From the above charts we observe how the relationship between presidential popularity and personal finances is somehow linear.  
Furthermore, from the boxplot we observe how the tenure appears to have a significant impact in the incumbent vote.


Following we output the results of the following linear model: 
$$IncVote = \beta_{0}+\beta_{1}PresPop+\beta_{2}PersFin+\beta_{3}Tenure+\varepsilon$$
```{r results="asis"}
stargazer(lm.holbrook, type = "html" )
```

As we can observe, only the coefficients PersFin and Tenure are statistically significant. This results is not surprising as PresPop and PersFin had a high correlation, meaning that somehow there were explaining the same variability.  
Following we analyze the residuals and verify the model assumptions.

```{r fig.height=12}
autoplot(lm.holbrook, which = 1:6, ncol = 2, label.size = 3)
```

As we can observe there the assumptions of normality and homoscedasticity are not satisfied.

```{r eval=FALSE}
lm.holbrook %>% 
  augment() %>% 
  mutate(stud.res = .resid / (lm.holbrook.summ$sigma * sqrt(1 - hatvalues(lm.holbrook)))) %>%
  ggplot(aes(stud.res)) +
  geom_histogram(binwidth = 0.05)
```

Following we determine whether there are some influential or outliers observations. We will use the following cutoffs to identify these observations:  
1. Cutoff CooksD 4/13 = 0.3 
2. Cutoff hat values (2*4)/13 = 0.6  

```{r}
lm.holbrook %>% 
  augment() %>% 
  filter(.hat > 2 * 4 / 13 | .cooksd > 4 / 13)
```

### Proposed Model
```{r}
lm.holbrook2 <- lm(IncVote ~ PersFin + Tenure, data = holbrook)
lm.holbrook2.summ <- summary(lm.holbrook2)
```

```{r}
holbrook %>%
  ggplot(aes(PersFin, IncVote, color = Tenure)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "lm") +
  labs(title= "Relationship between Incumbent Vote and Personal Finance")
```

Following we output the results of the following linear model: 
$$IncVote = \beta_{0}+\beta_{1}PersFin+\beta_{2}Tenure+\varepsilon$$

```{r results="asis"}
stargazer(lm.holbrook2, type = "html" )
```

```{r fig.height=12}
autoplot(lm.holbrook2, which = 1:6, ncol = 2, label.size = 3)
```

```{r eval=FALSE}
lm.holbrook2 %>% 
  augment() %>% 
  mutate(stud.res = .resid / (lm.holbrook.summ$sigma * sqrt(1 - hatvalues(lm.holbrook)))) %>%
  ggplot(aes(stud.res)) +
  geom_histogram(binwidth = 0.5)
```

```{r}
lm.holbrook2 %>% 
  augment() %>% 
  filter(.hat > 2 * 4 / 13 | .cooksd > 4 / 13)
```

## Lockerbie Model

```{r}
lockerbie %>% 
  gather(IncVote, Income1, Income2, NextYr, Tenure, key = "Var", value = "Value") %>% 
  ggplot(aes(Year, Value)) + 
  geom_line() +
  facet_grid(Var ~ ., scales = "free")
```

```{r}
lockerbie %>% 
  gather(Income1, Income2, NextYr, Tenure, key = "Var", value = "Value") %>% 
  ggplot(aes(Value, IncVote)) + 
  geom_point() +
  facet_grid(. ~ Var, scales = "free")
```


```{r}
lm.lockerbie <- lm(IncVote ~ Income1 + Income2 + NextYr + Tenure, data = lockerbie)
summary(lm.lockerbie)
```

```{r}
autoplot(lm.lockerbie, which = 1:6, ncol = 2, label.size = 3)
```


```{r}
bgtest(lm.lockerbie, order = 5)
```
